---
title: "Theory of Change"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

## Purpose

The purpose of this file is to analyze the data collected from the survey and to present the findings in a clear and concise manner. The analysis will include descriptive statistics, visualizations, and any relevant insights that can be drawn from the data.

## Data Ingestion

```{r}

# Load necessary libraries

library(tidyverse)

library(readxl)

library(janitor)

# Read the survey data

# survey_data <- readxl::read_excel("data/Consumer Behavior towards Organic Food Comparative Study_ Poland vs Zimbabwe (Responses) (2) (1).xlsx")

survey_data <- readxl::read_excel("F:/playground/The-Theory-of-Cognitive-Affective-Mediation-and-The-Moderating-Hypothesis-of-Price-Barriers-/data/Consumer Behavior towards Organic Food Comparative Study_ Poland vs Zimbabwe (Responses) (2) (1).xlsx")

# View the structure of the data

survey_data |> 
  tibble()

```

## Data Cleaning

```{r}

dim(survey_data)

# survey_data |>

# filter(!is.na(select_your_language_wybierz_jezyk))|>

# count(select_your_language_wybierz_jezyk)

# Define the mapping as a named list

coalesce_map <- list(
  consent = c("Do you agree to participate in this study after reading the information above?",
              "Czy po zapoznaniu się z powyższymi informacjami wyraża Pan/Pani zgodę na udział w tym badaniu?"),
  purchased_organic_last_12m = c( "A1. Have you purchased or consumed any organic food products in the last 12 months?","A1. Czy w ciągu ostatnich 12 miesięcy kupował (a) Pan/Pani lub spożywał(a) produkty ekologiczne?"),
  age = c('A2. Age:', 'A2. Wiek:'),
  gender = c('A3. Gender', 'A3. Płeć:'),
  education = c('A4. Highest Educational Level Completed:', 'A4. Najwyższy ukończony poziom wykształcenia:'),
  household_econ = c('A5. How would you describe the economic situation of your household?', 'A5. Jak ocenia Pani/Pan sytuację ekonomiczną swojego gospodarstwa domowego?'),
  employment = c('A6. Current Employment Status:', 'A6. Obecny status zatrudnienia:'),
  country = c('A7. Country of Residence:', 'A7. Kraj zamieszkania:'),
  awareness_duration = c('B1. How long have you been aware of organic food products?', 'B1. Od jakiego czasu zna Pan/i produkty żywności ekologicznej?'),
  pct_income_food = c('B2. Approximately what percentage of your total monthly household income is spent on food?', 'B2. Jaki procent swojego całkowitego miesięcznego dochodu gospodarstwa domowego przeznacza Pan/i średnio na żywność?'),
  pct_foodbudget_organic = c('B3. Of your total food budget, what percentage is typically spent on organic products?', 'B3. Jaki procent swojego budżetu przeznaczonego na żywność wydaje Pan/i średnio na produkty ekologiczne?'),
  purchase_frequency = c('B4. How often do you typically purchase organic food?', 'B4. Jak często kupuje Pan/i żywność ekologiczną?'),
  area_population = c('B5. What is the population size of your area of residence?', 'B5. Jaka jest wielkość populacji Pana/i miejsca zamieszkania?'),
  buy_where = c('B6. Where do you most often buy organic food? (Select all that apply)', 'B6. Gdzie najczęściej kupuje Pan/i żywność ekologiczną? (Można wybrać kilka odpowiedzi)'),
  buy_for_whom = c('B7. For whom do you primarily buy organic food? (Select all that apply)', 'B7. Dla kogo głównie kupuje Pan/i żywność ekologiczną? (Można wybrać kilka odpowiedzi)'),
  identify_method = c('B8. What is your primary method for identifying organic products?', 'B8. Jaka jest Pana/i główna metoda identyfikacji produktów ekologicznych? (Proszę wybrać jedną odpowiedź)'),
  c1_organic_belief = c('Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C1. Organic food is healthier than conventional food.]', 'Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C1. Żywność ekologiczna jest zdrowsza niż konwencjonalna.]'),
  c2_harmful_chemicals_belief = c('Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C2. Avoiding pesticides and harmful chemicals is a major advantage of organic food.]', 'Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C2. Unikanie pestycydów i szkodliwych chemikaliów jest główną zaletą żywności ekologicznej.]'),
  c3_environment_belief = c('Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C3. Organic farming is better for the environment.]', 'Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C3. Rolnictwo ekologiczne jest lepsze dla środowiska]'),
  c4_ethics_belief = c('Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C4. Organic farming ensures better ethical treatment of animals.]', 'Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C4. Rolnictwo ekologiczne zapewnia bardziej etyczne traktowanie zwierząt.]'),
  d1_organic_emotions = c( "Please indicate your level of agreement with the following statements based on your feelings. [D1. Organic food tastes better than non-organic food.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D1. Żywność ekologiczna smakuje lepiej niż konwencjonalna.]"),
  d2_assurance_emotion = c( "Please indicate your level of agreement with the following statements based on your feelings. [D2. I feel confident and reassured when I see an official organic certification label.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D2. Czuję się pewnie i spokojnie, gdy widzę logo ekożywności]"),
  d3_satisfaction_emotion = c( "Please indicate your level of agreement with the following statements based on your feelings. [D3. Buying organic food gives me a sense of personal satisfaction.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D3. Zakup żywności ekologicznej daje mi poczucie osobistej satysfakcji.]"),
  d4_health_worry_emotion = c( "Please indicate your level of agreement with the following statements based on your feelings. [D4. I feel worried about the long-term health effects of consuming conventional food.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D4. Czuję się zaniepokojony/a długoterminowymi skutkami zdrowotnymi spożywania konwencjonalnej żywności.]"),
  d5_trust_emotion = c( "Please indicate your level of agreement with the following statements based on your feelings. [D5. I feel a stronger sense of trust when I buy from a local farmer I know than from a certified supermarket product.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D5. Czuję większe zaufanie, gdy kupuję od znanego mi lokalnego rolnika niż do certyfikowanego produktu z supermarketu.]"),
  cert_ZW = c("C6. Zimbabwe Organic Certification Logo\r\nHow familiar are you with this logo?", "C6. Czy rozpoznaje Pan/i powyższe logo certyfikacji ekologicznej Zimbabwe?"),
  cert_familiar = c("C7. How important is seeing an official organic certification logo when you shop?", "C7. Jak ważne jest dla Pana/Pani oficjalne logo certyfikatu ekologicznego podczas zakupów?", "C7. How important is seeing an official organic certification logo when you shop? 2", "C7. Jak ważne jest dla Pana/Pani widok oficjalnego logo certyfikatu ekologicznego podczas zakupów?"),
  price_perception = c("E3. “Organic food is too expensive for my budget.”", "E3. Żywność ekologiczna jest dla mnie zbyt droga."),
  f1_intent = c("F1. Compared to now, do you intend to increase your consumption of organic food in the future?", "F1. W porównaniu do obecnej sytuacji, czy zamierza Pan/Pani zwiększyć w przyszłości spożycie żywności ekologicznej?\r\n")
  )

# survey_data |> dplyr::select(starts_with("A5")) |> names()

# Coalesce columns

for (canonical in names(coalesce_map)) {
  cols <- coalesce_map[[canonical]]
  survey_data[[canonical]] <- coalesce(!!!survey_data[cols])
  }

# Optionally, remove the original columns

# survey_data <- survey_data[, !(names(survey_data) %in% unlist(coalesce_map))]

# Get only Polish and Zimbabwean respondents

survey_data_clean <- survey_data |>
  clean_names() |>
  filter(country %in% c("Poland", "Polska", "Zimbabwe")) |>
  mutate(country = case_when(country %in% c("Poland", "Polska") ~ "Poland",
                             country == "Zimbabwe" ~ "Zimbabwe",
                             TRUE ~ country))

# Clean gender variable

survey_data_clean <- survey_data_clean |>
  mutate(gender_clean = case_when(
    str_to_lower(gender) %in% c("female", "kobieta", "woman") ~ "Female",
    str_to_lower(gender) %in% c("male", "mężczyzna", "mezczyzna", "man") ~ "Male",
    TRUE ~ gender))


# select only 100 polish and 100 zimbabwean respondents
set.seed(123) # For reproducibility

survey_data_sample <- survey_data_clean |>
  group_by(country) |>
  slice_sample(n = 100) |>
  ungroup()

```



```{r}

required_pkgs <- c(
  "tidyverse", # data wrangling & visualisation
  "lavaan", # SEM & mediation (Ha1, Ha2)
  "semPlot", # SEM path diagrams
  "mediation", # causal mediation analysis
  "lme4", # mixed models
  "interactions",# moderation plots (Ha3)
  "sandwich", # robust SEs
  "lmtest", # coefficient tests
  "psych", # composite reliability, alpha
  "broom", # tidy model outputs
  "modelsummary",# publication tables
  "car" # VIF, Anova
)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[,"Package"]]
if (length(to_install)) install.packages(to_install, dependencies = TRUE)
invisible(lapply(required_pkgs, library, character.only = TRUE))

```

## The Theory of Cognitive-Affective Mediation

`Alternative (H₁)`: Affective emotions are positively impacted by cognitive beliefs, which raise the frequency of purchases.

`Null (H₀)`: Affective sentiments do not enhance the frequency of purchases, while cognitive ideas have no effect on affective feelings.

```{r}

# ==================================================================

# Define mapping for purchase frequency

purchase_freq_map <- c(
  "daily" = 5,
  "codziennie" = 5,
  "a few times a week" = 4,
  "kilka razy w tygodniu" = 4,
  "once a week" = 3,
  "raz w tygodniu" = 3,
  "2–3 times a month" = 2,
  "2-3 razy w miesiącu" = 2,
  "once a month" = 2,
  "raz w miesiącu" = 2,
  "less than once a month / occasionally" = 1,
  "rzadziej niż raz w miesiącu / sporadycznie" = 1
)

# Clean and harmonize purchase frequency

survey_data_clean_df <- survey_data_sample |>
  mutate(
    purchase_frequency_clean = str_to_lower(str_trim(purchase_frequency)),
    purchase_frequency_clean = str_replace_all(purchase_frequency_clean, "–", "-"),
    purchase_frequency_rank = purchase_freq_map[purchase_frequency_clean]
)

# survey_data_clean_df %>% count(purchase_frequency_rank) |> print(n = Inf)

# ==================================================================

# library(dplyr)

df_ha1 <- survey_data_clean_df %>%
  dplyr::select(country, purchase_frequency_rank,
                c1_organic_belief:c4_ethics_belief,
                d1_organic_emotions:d5_trust_emotion)

# df_ha1 %>% dplyr::select(purchase_frequency_rank)

# # Count values per cur_column

# for (col in names(df_ha1)) {

# cat("n---", col, "value counts ---n")

# print(table(df_ha1[[col]], useNA = "ifany"))

# }

# Define Likert mapping (English + Polish)

likert_map <- c(
  "strongly disagree" = 1,
  "disagree" = 2,
  "neutral" = 3,
  "agree" = 4,
  "strongly agree" = 5,
  "zdecydowanie nie zgadzam się" = 1,
  "zdecydowanie się nie zgadzam" = 1,
  "nie zgadzam się" = 2,
  "raczej się nie zgadzam" = 2,"neutralny" = 3,
  "nie mam opinii" = 3,
  "ani się zgadzam, ani się nie zgadzam" = 3,
  "zgadzam się" = 4,
  "raczej się zgadzam" = 4,
  "strongly agree" = 5,
  "zdecydowanie zgadzam się" = 5,
  "zdecydowanie się zgadzam" = 5)

# List columns to clean

belief_emotion_cols <- c(
  "c1_organic_belief", "c2_harmful_chemicals_belief", "c3_environment_belief", "c4_ethics_belief",
  "d1_organic_emotions", "d2_assurance_emotion", "d3_satisfaction_emotion", "d4_health_worry_emotion",
  "d5_trust_emotion")

# Clean and convert to numeric

df_ha1 <- survey_data_clean_df %>%
  dplyr::select(country, purchase_frequency_rank,
                c1_organic_belief, c2_harmful_chemicals_belief, 
                c3_environment_belief, c4_ethics_belief,
                d1_organic_emotions, d2_assurance_emotion, 
                d3_satisfaction_emotion, d4_health_worry_emotion, 
                d5_trust_emotion) %>%
  mutate(across(all_of(belief_emotion_cols), 
                ~likert_map[str_to_lower(str_trim(.))],.names = "{.col}_num")) %>%
  mutate(purchase_frequency_rank = case_when(
    is.na(purchase_frequency_rank) ~ NA_integer_,
    purchase_frequency_rank == 'Invalid Number' ~ NA_integer_,
    TRUE ~ as.numeric(purchase_frequency_rank)))

# Creating composite scores for Cognitive and Affective constructs

df_ha1_df <- df_ha1 %>%
  mutate(
    Cognitive = rowMeans(cbind(
      c1_organic_belief_num, c2_harmful_chemicals_belief_num,
      c3_environment_belief_num, c4_ethics_belief_num), na.rm = TRUE),
    Affective = rowMeans(cbind(
      d1_organic_emotions_num, d2_assurance_emotion_num,
      d3_satisfaction_emotion_num, d4_health_worry_emotion_num #, d5_trust_emotion_num
      ), na.rm = TRUE))

# Baron-Kenny mediation

library(lmtest)
library(sandwich)

# Filter for complete cases

mediation_data <- df_ha1_df %>%
  dplyr::select(Cognitive, Affective, purchase_frequency_rank) %>%
  drop_na()

# Step 1: Cognitive → Purchase Frequency
# bk1 <- lm(purchase_frequency_rank ~ Cognitive, data = df_ha1_df)

bk1 <- lm(purchase_frequency_rank ~ Cognitive, data = mediation_data)
coeftest(bk1, vcov = vcovHC(bk1, "HC3"))

# Step 2: Cognitive → Affective
# bk2 <- lm(Affective ~ Cognitive, data = df_ha1_df)
bk2 <- lm(Affective ~ Cognitive, data = mediation_data)
coeftest(bk2, vcov = vcovHC(bk2, "HC3"))

# Step 3: Cognitive + Affective → Purchase Frequency
# bk3 <- lm(purchase_frequency_rank ~ Cognitive + Affective, data = df_ha1_df)
bk3 <- lm(purchase_frequency_rank ~ Cognitive + Affective, data = mediation_data)
coeftest(bk3, vcov = vcovHC(bk3, "HC3"))

# Fit models on mediation_data Mediation analysis

library(mediation)
set.seed(2025)
med <- mediation::mediate(
  model.m = bk2,
  model.y = bk3,
  treat = "Cognitive",
  mediator = "Affective",
  boot = TRUE,
  sims = 2000,
  boot.ci.type = "bca")

summary(med)

# structural Equation Modelling

#+ d5_trust_emotion_num

library(lavaan)

ha1_sem <- '
Cog =~ c1_organic_belief_num + c2_harmful_chemicals_belief_num + c3_environment_belief_num + c4_ethics_belief_num
Aff =~ d1_organic_emotions_num + d2_assurance_emotion_num + d3_satisfaction_emotion_num + d4_health_worry_emotion_num

Aff ~ a * Cog
purchase_frequency_rank ~ b * Aff + cp * Cog
indirect := a * b
total := cp + (a * b)
'

# Prepare SEM data with all item columns

sem_data <- df_ha1_df %>%
  dplyr::select(
    c1_organic_belief_num, c2_harmful_chemicals_belief_num,
    c3_environment_belief_num, c4_ethics_belief_num,
    d1_organic_emotions_num, d2_assurance_emotion_num,
    d3_satisfaction_emotion_num, d4_health_worry_emotion_num, #d5_trust_emotion_num
    Cognitive, Affective, purchase_frequency_rank
    ) %>%
  drop_na()

# Fit SEM model

ha1_fit <- sem(ha1_sem,
               data = sem_data,
               estimator = "ML",
               se = "bootstrap",
               bootstrap = 2000,
               missing = "ml")

library(semPlot)
semPaths(ha1_fit, what = "std",
         layout = "circle",
         edge.label.cex = 1.2, nCharNodes = 0,
         sizeMan = 8,
         sizeLat = 10, title = FALSE)

# Fitted Measures

fitMeasures(ha1_fit, 
            c("cfi", "tli", "rmsea", "srmr", "aic", "bic"))

summary(ha1_fit, standardized = TRUE, ci = TRUE)

# How to restructure?
modificationIndices(ha1_fit, sort = TRUE, maximum.number = 10)

```

## The Theory of the Certification Trust Chain

`Alternative ($H_1$)`: Purchase likelihood is increased by certification familiarity, which then boosts confidence after the purchase.
`Null ($H_0$)`: buy experience does not boost post-buy confidence, and certification familiarity does not raise purchase likelihood.



```{r}

# purchase likelihood (F1. "I intend to increase my purchase of organic food in the next 6 months."),

# certification familiarity (C5. EU Organic Logo (Used in Poland and the EU) How familiar are you with this logo? & "C6. Zimbabwe Organic Certification Logo How familiar are you with this logo?"

# confidence after purchase (D2: "I feel confident and reassured when I see an official organic certification label.")

df_ha2_df <- survey_data_sample |>
  dplyr::rename(
    c5_eu_logo_familiarity= c5_eu_organic_logo_used_in_poland_and_the_eu_how_familiar_are_you_with_this_logo,
    zim_logo_familiarity = cert_zw,
    organic_logo_importance = cert_familiar) |>
  dplyr::select(f1_intent, zim_logo_familiarity, organic_logo_importance, c5_eu_logo_familiarity,
                d2_assurance_emotion)



f1_intent_map <- c(
  "No, I intend to decrease it" = 1,
  "I’m not sure" = 2,
  "No, I will maintain the same level" = 3,
  "Pozostanie na tym samym poziomie" = 3,
  "Yes, slightly" = 4,
  "Raczej tak" = 4,
  "Yes, significantly" = 5,
  "Zdecydowanie tak" = 5)

# Harmonise

df_ha2_df$f1_intent_num <- f1_intent_map[str_trim(df_ha2_df$f1_intent)]

# Logo harmonisation

logo_fam_map <- c(
  "I do not recognize this logo at all" = 1,
  "Nie, nie rozpoznaję tego logo" = 1,
  "I have seen this logo but don’t know much about it" = 2,
  "Widziałem/am je, ale niewiele o nim wiem" = 2,
  "I recognize this logo and know what it means" = 3,
  "Znam to logo i wiem, co oznacza" = 3)

df_ha2_df$zim_logo_familiarity_num <- logo_fam_map[str_trim(df_ha2_df$zim_logo_familiarity)]
df_ha2_df$c5_eu_logo_familiarity_num <- logo_fam_map[str_trim(df_ha2_df$c5_eu_logo_familiarity)]

# Logo importance

logo_importance_map <- c(
  "Not important - It doesn't affect my purchasing decisions" = 1,
  "Nieważne - Nie wpływa na moje decyzje zakupowe" = 1,
  "I don't pay attention to certification logos" = 1,
  "Nie zwracam uwagi na logo certyfikatów" = 1,
  "Somewhat important - It influences my choice, but I sometimes buy uncertified" = 2,
  "Dość ważne - Wpływa na mój wybór, ale czasem kupuję produkty niecertyfikowane" = 2,
  "Very important - I look for it and usually only buy certified products" = 3)

df_ha2_df$organic_logo_importance_num <- logo_importance_map[str_trim(df_ha2_df$organic_logo_importance)]

# Mediation analysis

# Purchase likelihood (f1_intent) is increased by certification familiarity (Zim logo familiarity and EU logo familiarity), which then boosts confidence after the purchase (Organic food importance).

# Filter for complete cases

mediation_data_2 <- df_ha2_df %>%
  dplyr::select(f1_intent_num, c5_eu_logo_familiarity_num, zim_logo_familiarity_num,
                organic_logo_importance_num) %>%
  drop_na()

# Step 1: Certification familiarity → Purchase likelihood

bk1_2 <- lm(f1_intent_num ~ c5_eu_logo_familiarity_num + zim_logo_familiarity_num, data = mediation_data_2)
coeftest(bk1_2, vcov = vcovHC(bk1_2, "HC3"))

# Step 2: Certification familiarity → Organic logo importance

# bk2 <- lm(Affective ~ Cognitive, data = df_ha1_df)

bk2_2 <- lm(organic_logo_importance_num ~ c5_eu_logo_familiarity_num + zim_logo_familiarity_num, data = mediation_data_2)
coeftest(bk2_2, vcov = vcovHC(bk2_2, "HC3"))

# Step 3: Certification familiarity + Organic logo importance → Purchase likelihood

bk3_2 <- lm(f1_intent_num ~ c5_eu_logo_familiarity_num + zim_logo_familiarity_num + organic_logo_importance_num, data = mediation_data_2)
coeftest(bk3_2, vcov = vcovHC(bk3_2, "HC3"))

# Fit models on mediation_data Mediation analysis

set.seed(2025)
med_2 <- mediation::mediate(
  model.m = bk2_2,
  model.y = bk3_2,
  treat = "c5_eu_logo_familiarity_num",
  mediator = "organic_logo_importance_num",
  boot = TRUE,
  sims = 2000,
  boot.ci.type = "bca")

summary(med_2)

```

## The Moderating Hypothesis of Price Barriers


`Alternative ($H_1$)`: The positive correlation between affective sentiments and purchase intentions is weakened by price perception, particularly for consumers with lower incomes.

`Null ($H_0$)`: Regardless of economic level, the link between affective emotions and purchase intentions is not weakened by price perception.

```{r}

# =============================================================================

# Ha3: PRICE BARRIERS MODERATION
# Affective Emotions × Price Perception → Purchase Intention (F1)
# Moderated by Income Group (SES proxy)
# =============================================================================

# e3_zywnosc_ekologiczna_jest_dla_mnie_zbyt_droga
# e3_organic_food_is_too_expensive_for_my_budget
# Define Likert mapping (English + Polish)

likert_map <- c(
  "strongly disagree" = 1,
  "disagree" = 2,
  "neutral" = 3,
  "agree" = 4,
  "strongly agree" = 5,
  "zdecydowanie nie zgadzam się" = 1,
  "zdecydowanie się nie zgadzam" = 1,
  "nie zgadzam się" = 2,
  "raczej się nie zgadzam" = 2,
  "neutralny" = 3,
  "nie mam opinii" = 3,
  "ani się zgadzam, ani się nie zgadzam" = 3,
  "zgadzam się" = 4,
  "raczej się zgadzam" = 4,
  "strongly agree" = 5,
  "zdecydowanie zgadzam się" = 5,
  "zdecydowanie się zgadzam" = 5)

affective_emotions_list <- c(
  "d1_organic_emotions", "d2_assurance_emotion",
  "d3_satisfaction_emotion", "d4_health_worry_emotion" #, "d5_trust_emotion"
)

# Harmonize and numerically code 'employment' column

# Mapping both Polish and English values to numeric codes and English labels

employment_map <- c(
  'Employed full-time' = 1,
  'Praca na pełny etat' = 1,
  'Employed part-time' = 2,
  'Praca w niepełnym wymiarze czasu' = 2,
  'Self-employed / Freelancer' = 3,
  'Bezrobotna/ny' = 4,
  'Unemployed' = 4,
  'Retired' = 5,
  'Student' = 6,
  'Homemaker' = 7,
  'Osoba prowadząca gospodarstwo domowe' = 7
  )

employment_label_map <- c(
  'Employed full-time' = 'Employed full-time',
  'Praca na pełny etat' = 'Employed full-time',
  'Employed part-time' = 'Employed part-time',
  'Praca w niepełnym wymiarze czasu' = 'Employed part-time',
  'Self-employed / Freelancer' = 'Self-employed / Freelancer',
  'Bezrobotna/ny' = 'Unemployed',
  'Unemployed' = 'Unemployed',
  'Retired' = 'Retired',
  'Student' = 'Student',
  'Homemaker' = 'Homemaker',
  'Osoba prowadząca gospodarstwo domowe' = 'Homemaker')

survey_data_sample <- survey_data_sample |> 
  mutate(
    employment_num = employment_map[employment],
    employment_label = employment_label_map[employment])

# Clean and convert to numeric

df_ha3 <- survey_data_sample %>%
  dplyr::select(f1_intent, price_perception, employment_num,
                all_of(affective_emotions_list)) %>%
  mutate(across(all_of(affective_emotions_list),
                ~ likert_map[str_to_lower(str_trim(.))],
                .names = "{.col}_num"))

# Creating composite scores for Cognitive and Affective constructs

df_ha3_df <- df_ha3 %>%
  mutate(Affective = rowMeans(cbind(
    d1_organic_emotions_num, d2_assurance_emotion_num,
    d3_satisfaction_emotion_num, d4_health_worry_emotion_num #, d5_trust_emotion_num
    ), na.rm = TRUE))

# Harmonize and numerically code 'price_perception' column
# Mapping: Strongly disagree = 1, Disagree = 2, Neutral = 3, Agree = 4, Strongly agree = 5
# Polish equivalents: 'Zdecydowanie się nie zgadzam' = 1, 'Nie zgadzam się' = 2, 'Ani się zgadzam, ani nie' = 3, 'Zgadzam się' = 4, 'Zdecydowanie się zgadzam' = 5
# For multiple answers, take the mean of mapped values

price_perception_map <- c(
  'Strongly disagree' = 1,
  'Disagree' = 2,
  'Neutral' = 3,
  'Agree' = 4,
  'Strongly agree' = 5,
  'Zdecydowanie się nie zgadzam' = 1,
  'Nie zgadzam się' = 2,
  'Ani się zgadzam, ani nie' = 3,
  'Zgadzam się' = 4,
  'Zdecydowanie się zgadzam' = 5)

df_ha3_df <- df_ha3_df |> 
  mutate(
    price_perception_num = sapply(strsplit(price_perception, ','), 
                                  function(x) {x <- trimws(x)
                                  vals <- price_perception_map[x]
                                  as.numeric(vals[1]) # Use only the first mapped value
                                  }))



# Check the result

# df_ha3_df |> count(household_econ, household_econ_num)

f1_intent_map <- c(
  "No, I intend to decrease it" = 1,
  "I’m not sure" = 2,
  "No, I will maintain the same level" = 3,
  "Pozostanie na tym samym poziomie" = 3,
  "Yes, slightly" = 4,
  "Raczej tak" = 4,
  "Yes, significantly" = 5,
  "Zdecydowanie tak" = 5)

# Harmonise

df_ha3_df$f1_intent_num <- f1_intent_map[str_trim(df_ha3_df$f1_intent)]

# Check the result

# df_ha3_df |> count(price_perception, price_perception_num)

# survey_data_sample |> count(employment)

# Check the result

# survey_data_sample |> count(employment, employment_num, employment_label)

# Droping rows with NA in key variables

df_ha3_df <- df_ha3_df |> drop_na(f1_intent, Affective, price_perception_num, employment_num)

# df_ha3_df <- df_ha3_df |>
#   mutate(
#     Affective_c = Affective - mean(Affective, na.rm = TRUE),
#     E3_price_c = price_perception_num - mean(price_perception_num, na.rm = TRUE))

# ── Ha3-A: Base 2-way moderation ──────────────────────────────────────────────

ha3_base <- lm(f1_intent_num ~ Affective * price_perception_num, data = df_ha3_df)
cat("n--- Ha3 Base: Affective × Price → F1_intent ---n")
print(coeftest(ha3_base, vcov = vcovHC(ha3_base, "HC3")))
cat(sprintf("R² = %.4f", summary(ha3_base)$r.squared))

# ── Ha3-B: 3-way with income group ────────────────────────────────────────────

ha3_3way <- lm(f1_intent_num ~ Affective * price_perception_num * employment_num, data = df_ha3_df)
cat("n--- Ha3 3-Way: Affective × Price × Employment → F1_intent ---n")
print(coeftest(ha3_3way, vcov = vcovHC(ha3_3way, "HC3")))
cat(sprintf("R² = %.4f ΔR² = %.4fn",
            summary(ha3_3way)$r.squared,
            summary(ha3_3way)$r.squared - summary(ha3_base)$r.squared))

# ── Ha3-C: Model comparison ─────`───────────────────────────────────────────────

cat("n--- Ha3: ANOVA model comparison ---n")
print(anova(ha3_base, ha3_3way))

```

## Hypothesis of Country-Specific Demographic Drivers

`Alternative ($H_1$)`: In both nations, socioeconomic position has a favorable impact on cognitive beliefs and purchase frequency, but it does so through distinct mechanisms—directly in Zimbabwe and indirectly through education and urban residency in Poland.

`Null ($H_0$)`: Neither country's cognitive views nor purchase frequency are significantly impacted by socioeconomic position.

Socioeconomic position (SEP) is defined by:
1. A5 — Household Economic Situation
2. A4 — Highest Educational Level Completed
3. A6 — Current Employment Status

```{r}
# =============================================================================
# Ha4: COUNTRY-SPECIFIC DEMOGRAPHIC DRIVERS
# Zimbabwe: SES → Cognitive / Purchase directly
# Poland: SES → (Educ + Urban) → Cognitive → Purchase (indirect pathway)
# =============================================================================


# Harmonize and numerically code 'household_econ' column

# # Mapping: Poor = 1, Fair/Average/Przeciętna = 2, Good/Dobra = 3, Very good = 4

household_econ_map <- c(
  'Poor' = 1,
  'Dobra' = 3,
  'Good' = 3,
  'Fair/Average' = 2,
  'Przeciętna' = 2,
  'Very good' = 4)


# zim_data <- survey_data_sample %>% 
#   filter(country == "Zimbabwe") %>%
#   mutate(household_econ_num = household_econ_map[household_econ]) %>% names()
#   select(household_econ, education, 
#          Cognitive, purchase_frequency_rank)


```

