---
title: "final_analysis"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
required_pkgs <- c(

  "tidyverse",    # data wrangling & visualisation

  "readxl",       # Excel ingestion

  "janitor",      # clean_names()

  "lavaan",       # SEM & CFA

  "semPlot",      # SEM path diagrams

  "mediation",    # bootstrapped causal mediation

  "lmtest",       # coeftest()

  "sandwich",     # HC3 robust standard errors

  "interactions", # interact_plot(), johnson_neyman()

  "psych",        # alpha(), describe()

  "broom",        # tidy model outputs

  "modelsummary", # publication-ready tables

  "car"           # vif(), Anova()

)

to_install <- required_pkgs[!required_pkgs %in% installed.packages()[, "Package"]]

if (length(to_install)) install.packages(to_install, dependencies = TRUE)

invisible(lapply(required_pkgs, library, character.only = TRUE))

# Resolve common namespace conflicts explicitly

select  <- dplyr::select
filter  <- dplyr::filter
recode  <- dplyr::recode
mediate <- mediation::mediate


```

```{r}
LIKERTS <- c(
  # English
  "strongly disagree"                         = 1L,
  "disagree"                                  = 2L,
  "neutral"                                   = 3L,
  "agree"                                     = 4L,
  "strongly agree"                            = 5L,
  # Polish variants
  "zdecydowanie nie zgadzam się"              = 1L,
  "zdecydowanie się nie zgadzam"              = 1L,
  "nie zgadzam się"                           = 2L,
  "raczej się nie zgadzam"                    = 2L,
  "neutralny"                                 = 3L,
  "nie mam opinii"                            = 3L,
  "ani się zgadzam, ani się nie zgadzam"      = 3L,
  "ani się zgadzam, ani nie"                  = 3L,
  "zgadzam się"                               = 4L,
  "raczej się zgadzam"                        = 4L,
  "zdecydowanie zgadzam się"                  = 5L,
  "zdecydowanie się zgadzam"                  = 5L
)

# ── 1.2  Purchase frequency ───────────────────────────────────────────────────

PURCHASE_FREQ <- c(
  "less than once a month / occasionally"               = 1L,
  "rzadziej niż raz w miesiącu / sporadycznie"          = 1L,
  "once a month"                                        = 2L,
  "raz w miesiącu"                                      = 2L,
  "2–3 times a month"                                   = 3L,
  "2-3 razy w miesiącu"                                 = 3L,
  "once a week"                                         = 4L,
  "raz w tygodniu"                                      = 4L,
  "a few times a week"                                  = 5L,
  "kilka razy w tygodniu"                               = 5L,
  "daily"                                               = 6L,
  "codziennie"                                          = 6L
)

# ── 1.3  Purchase intention (F1) ──────────────────────────────────────────────

F1_INTENT <- c(
  "no, i intend to decrease it"      = 1L,
  "i'm not sure"                     = 2L,
  "no, i will maintain the same level" = 3L,
  "pozostanie na tym samym poziomie" = 3L,
  "yes, slightly"                    = 4L,
  "raczej tak"                       = 4L,
  "yes, significantly"               = 5L,
  "zdecydowanie tak"                 = 5L
)



# ── 1.4  Certification logo familiarity ───────────────────────────────────────

CERT_FAM <- c(
  "i do not recognize this logo at all"              = 1L,
  "nie, nie rozpoznaję tego logo"                    = 1L,
  "i have seen this logo but don't know much about it" = 2L,
  "widziałem/am je, ale niewiele o nim wiem"         = 2L,
  "i recognize this logo and know what it means"     = 3L,
  "znam to logo i wiem, co oznacza"                  = 3L
)




# ── 1.5  Certification / logo importance ─────────────────────────────────────

CERT_IMP <- c(
  "not important - it doesn't affect my purchasing decisions" = 1L,
  "nieważne - nie wpływa na moje decyzje zakupowe"           = 1L,
  "i don't pay attention to certification logos"             = 1L,
  "nie zwracam uwagi na logo certyfikatów"                   = 1L,
  "somewhat important - it influences my choice, but i sometimes buy uncertified" = 2L,
  "dość ważne - wpływa na mój wybór, ale czasem kupuję produkty niecertyfikowane" = 2L,
  "very important - i look for it and usually only buy certified products"         = 3L
)



# ── 1.6  Household economic situation (SES proxy) ────────────────────────────

HOUSEHOLD_ECON <- c(
  "poor"         = 1L,
  "fair/average" = 2L,
  "przeciętna"   = 2L,
  "good"         = 3L,
  "dobra"        = 3L,
  "bardzo dobra" = 4L,
  "very good"    = 4L
)


# ── 1.7  Employment status ────────────────────────────────────────────────────

# EMPLOYMENT <- c(
#   "Employed full-time"                        = 1L,
#   "Praca na pełny etat"                       = 1L,
#   "Employed part-time"                        = 2L,
#   "Praca w niepełnym wymiarze czasu"          = 2L,
#   "self-employed / Freelancer"                = 3L,
#   "Unemployed"                                = 4L,
#   "Bezrobotna/ny"                             = 4L,
#   "Retired"                                   = 5L,
#   "Student"                                   = 6L,
#   "Homemaker"                                 = 7L,
#   "Osoba prowadząca gospodarstwo domowe"      = 7L
# )

EMPLOYMENT <- c(
  "employed full-time"                   = 5L,
  "praca na pełny etat"                  = 5L,
  "self-employed / freelancer"           = 5L,
  "employed part-time"                   = 4L,
  "praca w niepełnym wymiarze czasu"     = 4L,
  "student"                              = 3L,
  "homemaker"                            = 3L,
  "osoba prowadząca gospodarstwo domowe" = 3L,
  "retired"                              = 3L,
  "unemployed"                           = 2L,
  "bezrobotna/ny"                        = 2L,
  "NA"                                   = 1L
)





# ── 1.8  Education ────────────────────────────────────────────────────────────

EDUCATION <- c(
  # Level 1: Primary
  "primary or Less"                            = 1L,
  "podstawowe"                                 = 1L,
  
  # Level 2: Vocational (Lower Secondary/Trade)
  "szkoła zawodowa"                            = 2L, 
  
  # Level 3: Secondary / High School
  "secondary/high school"                      = 3L,
  "średnie / liceum / technikum"               = 3L,
  "technical/vocational training"              = 3L, # Technikum is often here
  "college"                                    = 3L,
  
  # Level 4: Bachelor's / Engineer
  "bachelor's degree"                          = 4L,
  "wykształcenie wyższe (licencjat/inżynier)"  = 4L,
  
  # Level 5: Master's or PhD
  "master's degree or higher"                  = 5L,
  "postgraduate Degree (masters, phd)"         = 5L,
  "doctorate"                                  = 5L,
  "wykształcenie wyższe (magister lub wyższe)" = 5L,
  "magister lub wyższy"                        = 5L
)



```

```{r}
# ── 1.9  Colour palette (reused in all plots) ─────────────────────────────────

COL_COUNTRY <- c("Poland" = "#e41a1c", "Zimbabwe" = "#377eb8")

# ── 2.1  Read raw Excel ───────────────────────────────────────────────────────

# Update the path to point to your local copy of the Excel file.

raw <- readxl::read_excel(
  "data/Consumer Behavior towards Organic Food Comparative Study_ Poland vs Zimbabwe (Responses) (2) (1).xlsx"
)

```

```{r}
# ── 2.2  Bilingual column coalescing ─────────────────────────────────────────

# For each construct the survey has parallel English + Polish columns.

# coalesce() picks the first non-NA value, so bilingual responses map

# to a single canonical column.

coalesce_map <- list(
  consent = c("Do you agree to participate in this study after reading the information above?", "Czy po zapoznaniu się z powyższymi informacjami wyraża Pan/Pani zgodę na udział w tym badaniu?"),
  purchased_last_12m    = c("A1. Have you purchased or consumed any organic food products in the last 12 months?", "  A1. Czy w ciągu ostatnich 12 miesięcy kupował (a) Pan/Pani lub spożywał(a) produkty ekologiczne?  "),
  age = c("A2. Age:", "A2. Wiek:"),
  gender = c("A3. Gender", "A3. Płeć:"),
  education = c("A4. Highest Educational Level Completed:", "A4. Najwyższy ukończony poziom wykształcenia:"),
  household_econ = c("A5. How would you describe the economic situation of your household?", "A5. Jak ocenia Pani/Pan sytuację ekonomiczną swojego gospodarstwa domowego?"),
  employment = c("A6. Current Employment Status:", "A6. Obecny status zatrudnienia:"),
  country = c("A7. Country of Residence:", "A7. Kraj zamieszkania:"),
  awareness_duration = c("B1. How long have you been aware of organic food products?", "B1. Od jakiego czasu zna Pan/i produkty żywności ekologicznej?"),
  pct_income_food = c("B2. Approximately what percentage of your total monthly household income is spent on food?", "B2. Jaki procent swojego całkowitego miesięcznego dochodu gospodarstwa domowego przeznacza Pan/i średnio na żywność?"),
  pct_budget_organic = c("B3. Of your total food budget, what percentage is typically spent on organic products?", "B3. Jaki procent swojego budżetu przeznaczonego na żywność wydaje Pan/i średnio na produkty ekologiczne?"),
  purchase_frequency = c("B4. How often do you typically purchase organic food?", "B4. Jak często kupuje Pan/i żywność ekologiczną?"),
  area_population = c("B5. What is the population size of your area of residence?", "B5. Jaka jest wielkość populacji Pana/i miejsca zamieszkania?"),
  buy_where = c("B6. Where do you most often buy organic food? (Select all that apply)", "B6. Gdzie najczęściej kupuje Pan/i żywność ekologiczną? (Można wybrać kilka odpowiedzi)"),
  buy_for_whom = c("B7. For whom do you primarily buy organic food? (Select all that apply)", "B7. Dla kogo głównie kupuje Pan/i żywność ekologiczną? (Można wybrać kilka odpowiedzi)"),
  identify_method = c("B8. What is your primary method for identifying organic products?", "B8. Jaka jest Pana/i główna metoda identyfikacji produktów ekologicznych? (Proszę wybrać jedną odpowiedź)"),
  c1_organic_belief = c("Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C1. Organic food is healthier than conventional food.]", "Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C1. Żywność ekologiczna jest zdrowsza niż konwencjonalna.]"),
  c2_pesticides_belief  = c("Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C2. Avoiding pesticides and harmful chemicals is a major advantage of organic food.]", "Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C2. Unikanie pestycydów i szkodliwych chemikaliów jest główną zaletą żywności ekologicznej.]"),
  c3_environment_belief = c("Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C3. Organic farming is better for the environment.]", "Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C3. Rolnictwo ekologiczne jest lepsze dla środowiska]"),
  c4_ethics_belief = c("Part 1: Your Beliefs About Organic Food\r\nPlease indicate your level of agreement with the following statements based on your beliefs. [C4. Organic farming ensures better ethical treatment of animals.]",
"Część 1: Przekonania na temat żywności ekologicznej\r\nProszę wskazać stopień zgodności z poniższymi stwierdzeniami, opierając się na swoich przekonaniach. [C4. Rolnictwo ekologiczne zapewnia bardziej etyczne traktowanie zwierząt.]"),
cert_eu = c("C5. EU Organic Logo (Used in Poland and the EU)\r\nHow familiar are you with this logo?", "Proszę zapoznać się z poniższymi logo certyfikatów ekologicznych i odpowiedzieć na pytania.\r\nPICTURE OF EU ORGANIC LOGO\r\nrozpoznaje Pan/i powyższe logo ekologiczne UE?"),
cert_zw = c("C6. Zimbabwe Organic Certification Logo\r\nHow familiar are you with this logo?", "C6. Czy rozpoznaje Pan/i powyższe logo certyfikacji ekologicznej Zimbabwe?"),
cert_importance = c("C7. How important is seeing an official organic certification logo when you shop?", "C7. Jak ważne jest dla Pana/Pani oficjalne logo certyfikatu ekologicznego podczas zakupów?"),
d1_taste_emotion = c("Please indicate your level of agreement with the following statements based on your feelings. [D1. Organic food tastes better than non-organic food.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D1. Żywność ekologiczna smakuje lepiej niż konwencjonalna.]"),
d2_assurance_emotion  = c("Please indicate your level of agreement with the following statements based on your feelings. [D2. I feel confident and reassured when I see an official organic certification label.]","Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D2. Czuję się pewnie i spokojnie, gdy widzę logo ekożywności]"),

  d3_satisfaction_emotion = c("Please indicate your level of agreement with the following statements based on your feelings. [D3. Buying organic food gives me a sense of personal satisfaction.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D3. Zakup żywności ekologicznej daje mi poczucie osobistej satysfakcji.]"),
d4_health_worry = c("Please indicate your level of agreement with the following statements based on your feelings. [D4. I feel worried about the long-term health effects of consuming conventional food.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D4. Czuję się zaniepokojony/a długoterminowymi skutkami zdrowotnymi spożywania konwencjonalnej żywności.]"), 
d5_local_trust = c("Please indicate your level of agreement with the following statements based on your feelings. [D5. I feel a stronger sense of trust when I buy from a local farmer I know than from a certified supermarket product.]", "Proszę określić w jakim stopniu zgadza się Pan(i) z poniższymi stwierdzeniami, w oparciu o swoje odczucia. [D5. Czuję większe zaufanie, gdy kupuję od znanego mi lokalnego rolnika niż do certyfikowanego produktu z supermarketu.]"), 
price_perception = c("E3. \u201cOrganic food is too expensive for my budget.\u201d", "E3. Żywność ekologiczna jest dla mnie zbyt droga."),
f1_intent = c("F1. Compared to now, do you intend to increase your consumption of organic food in the future?", "F1. W porównaniu do obecnej sytuacji, czy zamierza Pan/Pani zwiększyć w przyszłości spożycie żywności ekologicznej?\n"))

# Apply coalescing: for each canonical name, combine bilingual columns

for (canonical in names(coalesce_map)) {
  cols <- coalesce_map[[canonical]]
  valid_cols  <- cols[cols %in% names(raw)]          # only keep cols that exist
  raw[[canonical]] <- do.call(coalesce, raw[valid_cols])
}
```

## Data Cleaning

```{r}


# ── 3.1  Helper: apply a named lookup vector (case-insensitive) ───────────────

# This replaces all repeated instances of map[str_to_lower(str_trim(x))].

apply_map <- function(x, lookup) {
  lookup[str_to_lower(str_trim(as.character(x)))]
}

# ── 3.2  Filter to Poland & Zimbabwe; standardise country labels ──────────────

df <- raw |>
  clean_names() |>
  filter(country %in% c("Poland", "Polska", "Zimbabwe")) |>
  mutate(
    country = if_else(country == "Polska", "Poland", country),
    country = factor(country, levels = c("Poland", "Zimbabwe"))
  )

# ── 3.3  Balanced sample: 100 per country ─────────────────────────────────────

set.seed(123)
df <- df |>
  group_by(country) |>
  slice_sample(n = 100) |>
  ungroup()

cat(sprintf("\nSample: %d total (%s)\n", nrow(df),
            paste(names(table(df$country)), table(df$country), sep = " = ",
                  collapse = " | ")))

# ── 3.4  Apply all lookup mappings ────────────────────────────────────────────

# Likert items: C1–C4 (cognitive beliefs) and D1–D5 (affective emotions)

cognitive_cols <- c("c1_organic_belief", "c2_pesticides_belief",
                    "c3_environment_belief", "c4_ethics_belief")

affective_cols <- c("d1_taste_emotion", "d2_assurance_emotion",
                    "d3_satisfaction_emotion", "d4_health_worry")

# Note: d5_local_trust excluded from composite (cross-loads; see MIs below)

likert_cols <- c(cognitive_cols, affective_cols)

df <- df |> 
  # Likert items → numeric
  mutate(across(all_of(likert_cols),
                ~ as.integer(apply_map(.x, LIKERTS)),
                .names = "{.col}_n")) |>
  # Purchase frequency → ordinal 1–6
  mutate(
    purchase_frequency_clean = str_replace_all(
      str_to_lower(str_trim(purchase_frequency)), "–", "-"),
    purchase_freq_n = as.integer(PURCHASE_FREQ[purchase_frequency_clean])
  ) |>
  # F1 purchase intention → ordinal 1–5
  mutate(
    f1_intent_n = as.integer(
      apply_map(f1_intent, F1_INTENT))
  ) |>
  # Certification familiarity (EU & ZW) → 1–3
  mutate(
    cert_eu_n    = as.integer(apply_map(cert_eu, CERT_FAM)),
    cert_zw_n    = as.integer(apply_map(cert_zw, CERT_FAM)),
    cert_imp_n   = as.integer(apply_map(cert_importance, CERT_IMP))
  ) |>
  # Socioeconomic proxies
  mutate(
    household_econ_n = as.integer(apply_map(household_econ, HOUSEHOLD_ECON)),
    employment_n     = as.integer(apply_map(employment, EMPLOYMENT)),
    education_n      = as.integer(apply_map(education, EDUCATION))
  ) |>
  # Price perception: take first response only (handles multi-select edge cases)
  mutate(
    price_perception_n = as.integer(
      apply_map(str_extract(price_perception, "^[^,]+"), LIKERTS))
  ) |>
  # Country-appropriate certification familiarity
  mutate(
    cert_familiar_n = if_else(country == "Poland", cert_eu_n, cert_zw_n)
  ) |>
  # Gender harmonisation
  mutate(
    gender_clean = case_when(
      str_to_lower(gender) %in% c("female", "kobieta", "woman") ~ "Female",
      str_to_lower(gender) %in% c("male",   "mężczyzna", "man") ~ "Male",
      TRUE ~ "Other / Prefer not to say"
    )
  )


```

```{r}
# ── 4.1  Cognitive and Affective composites ───────────────────────────────────

cog_items <- paste0(cognitive_cols, "_n")   # c1…c4_n
aff_items <- paste0(affective_cols, "_n")   # d1…d4_n  (d5 excluded)

df <- df |>
  mutate(
    Cognitive = rowMeans(across(all_of(cog_items)), na.rm = TRUE),
    Affective = rowMeans(across(all_of(aff_items)), na.rm = TRUE),
    # Mean-centred versions for moderation models (Ha3) to reduce multicollinearity
    Affective_c = Affective       - mean(Affective,         na.rm = TRUE),
    price_c = price_perception_n - mean(price_perception_n, na.rm = TRUE),
    # SEP composite (Ha4): mean of household SES, education, employment
    SEP = rowMeans(across(c(household_econ_n, education_n, employment_n)), na.rm = TRUE)
  )

# ── 4.2  Internal consistency ─────────────────────────────────────────────────

cat("\n--- Reliability ---\n")
cat("Cognitive Beliefs (C1–C4): α =",
    round(psych::alpha(df[, cog_items], na.rm = TRUE)$total$raw_alpha, 3), "\n")
cat("Affective Emotions (D1–D4): α =",
    round(psych::alpha(df[, aff_items], na.rm = TRUE)$total$raw_alpha, 3), "\n")

# ── 4.3  Descriptive overview ─────────────────────────────────────────────────

cat("\n--- Descriptive Statistics ---\n")
df |> 
  select(Cognitive, Affective, purchase_freq_n, cert_familiar_n,
         price_perception_n, f1_intent_n, SEP, education_n) |>
  psych::describe() |>
  round(2) |>
  print()





```

```{r}
# ── 5.1  Pretty section header ────────────────────────────────────────────────

section <- function(title) {
  bar <- strrep("=", 60)
  cat("\n\n", bar, "\n ", title, "\n", bar, "\n", sep = "")
}

# ── 5.2  Robust coefficient test with HC3 SEs ─────────────────────────────────

# Returns the coeftest object AND prints it with a label.

rob_test <- function(model, label = "") {
  if (nchar(label) > 0) cat("\n---", label, "---\n")
  ct <- coeftest(model, vcov = vcovHC(model, "HC3"))
  print(ct)
  invisible(ct)
}

# ── 5.3  Bootstrapped mediation (wrapper to keep sims consistent) ─────────────

run_mediation <- function(model_m, model_y, treat, mediator,
                           sims = 2000, seed = 2025) {
  set.seed(seed)
  med_out <- mediate(
    model.m       = model_m,
    model.y       = model_y,
    treat         = treat,
    mediator      = mediator,
    boot          = TRUE,
    sims          = sims,
    boot.ci.type  = "bca"
  )

  print(summary(med_out))
  invisible(med_out)
}

# ── 5.4  Sobel Z-test ─────────────────────────────────────────────────────────

sobel_test <- function(model_a, treat, model_b, mediator) {
  a    <- coef(model_a)[treat]
  b    <- coef(model_b)[mediator]
  se_a <- sqrt(vcov(model_a)[treat,    treat])
  se_b <- sqrt(vcov(model_b)[mediator, mediator])
  z    <- (a * b) / sqrt(b^2 * se_a^2 + a^2 * se_b^2)
  p    <- 2 * pnorm(-abs(z))
  cat(sprintf("\n--- Sobel Test ---\n  Z = %.4f  |  p = %.4f\n", z, p))
  invisible(list(z = z, p = p))
}

# ── 5.5  SEM fit summary ──────────────────────────────────────────────────────

sem_fit_summary <- function(fit, label = "") {
  if (nchar(label) > 0) cat("\n---", label, "---\n")
  cat("Fit indices:\n")
  print(fitMeasures(fit, c("cfi", "tli", "rmsea", "srmr", "aic", "bic")))
  invisible(fit)
}

# ── 5.6  Extract & print labelled SEM paths ───────────────────────────────────

sem_paths <- function(fit, labels) {
  parameterEstimates(fit, boot.ci.type = "bca.simple", ci = TRUE,
                     standardized = TRUE) |>
    filter(label %in% labels) |>
    select(label, est, se, z, pvalue, ci.lower, ci.upper, std.all) |>
    print()
}

# ── 5.7  H₀ rejection decision for indirect effect ───────────────────────────

decide_indirect <- function(fit, label = "indirect") {
  est <- parameterEstimates(fit, boot.ci.type = "bca.simple", ci = TRUE) |>
    filter(.data$label == !!label)
  cat(sprintf("\n>>> Decision (%s): est = %.3f [%.3f, %.3f]  p = %.4f\n",
              label, est$est, est$ci.lower, est$ci.upper, est$pvalue))
  if (est$ci.lower > 0 | est$ci.upper < 0) {
    cat("    REJECT H\u2080 \u2014 CI excludes zero.\n")
  } else {
    cat("    FAIL TO REJECT H\u2080 \u2014 CI includes zero.\n")
  }

  invisible(est)

}

# ── 5.8  Consistent ggplot2 theme ─────────────────────────────────────────────

theme_study <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(plot.title    = element_text(face = "bold"),
          plot.subtitle = element_text(colour = "grey40"),
          strip.text    = element_text(face = "bold"))
}

```

## **The Theory of Cognitive-Affective Mediation** 

`Alternative (H₁)`: Affective emotions are positively impacted by cognitive beliefs, which raise the frequency of purchases.

`Null (H₀)`: Affective sentiments do not enhance the frequency of purchases, while cognitive ideas have no effect on affective feelings.

```{r}

#| fig-width: 10
#| fig-height: 8
#| fig-align: "center"


# 6.  Ha1: THEORY OF COGNITIVE-AFFECTIVE MEDIATION
#
#  Path:  Cognitive Beliefs (C1–C4)
#            → Affective Emotions (D1–D4)   [a-path]
#            → Purchase Frequency           [b-path]
#  Also tests direct path Cognitive → Purchase (c'-path)
# 

ha1_data <- df |>
  select(country, all_of(c(cog_items, aff_items)),
         Cognitive, Affective, purchase_freq_n) |>
  drop_na()

cat(sprintf("\nHa1 n = %d\n", nrow(ha1_data)))

# ── 6.2  Baron-Kenny steps ────────────────────────────────────────────────────

# Step 1: total effect — Cognitive → Purchase

bk1 <- lm(purchase_freq_n ~ Cognitive, data = ha1_data)
rob_test(bk1, "Step 1 – Total effect: Cognitive → Purchase_Freq")

# Step 2: a-path — Cognitive → Affective

bk2 <- lm(Affective ~ Cognitive, data = ha1_data)
rob_test(bk2, "Step 2 – a-path: Cognitive → Affective")

# Step 3: b & c' paths — Cognitive + Affective → Purchase

bk3 <- lm(purchase_freq_n ~ Cognitive + Affective, data = ha1_data)
rob_test(bk3, "Step 3 – b & c' paths: Cognitive + Affective → Purchase_Freq")

# ── 6.3  Bootstrapped mediation ───────────────────────────────────────────────

cat("\n--- Ha1: Bootstrapped ACME / ADE / Total ---\n")
ha1_med <- run_mediation(bk2, bk3, treat = "Cognitive", mediator = "Affective")
# ── 6.4  SEM with CFA measurement model ──────────────────────────────────────

# Using ML estimator (MLR available but requires FIML; drop_na used above)

ha1_sem_spec <- '
  # Measurement model
  Cog =~ c1_organic_belief_n + c2_pesticides_belief_n +
         c3_environment_belief_n + c4_ethics_belief_n
  Aff =~ d1_taste_emotion_n + d2_assurance_emotion_n +
         d3_satisfaction_emotion_n + d4_health_worry_n

  # Structural model
  Aff           ~ a  * Cog
  purchase_freq_n ~ b  * Aff + cp * Cog
  # Derived effects
  indirect := a * b
  total    := cp + (a * b)
'

ha1_fit <- sem(ha1_sem_spec, data = ha1_data,
               estimator  = "ML",
               se         = "bootstrap",
               bootstrap  = 2000)

sem_fit_summary(ha1_fit, "Ha1 SEM Fit Indices")

cat("\n--- Ha1: Structural path estimates ---\n")
sem_paths(ha1_fit, c("a", "b", "cp", "indirect", "total"))

decide_indirect(ha1_fit, "indirect")

# ── 6.5  SEM path diagram ─────────────────────────────────────────────────────

semPlot::semPaths(ha1_fit, what = "std", layout = "circle",
                  edge.label.cex = 1.2, nCharNodes = 0,
                  sizeMan = 8, sizeLat = 10, title = FALSE)

# ── 6.6  Modification indices (for model improvement guidance) ────────────────

cat("\n--- Ha1: Top 10 modification indices ---\n")
modificationIndices(ha1_fit, sort. = TRUE, maximum.number = 10) |> 
  print()

# ── 6.7  Plot ─────────────────────────────────────────────────────────────────

ggplot(ha1_data, aes(x = Cognitive, y = purchase_freq_n)) +
  geom_jitter(aes(colour = Affective), alpha = 0.55, width = 0.04, height = 0.15) +
  geom_smooth(method = "lm", se = TRUE, colour = "#2171b5", linewidth = 1) +
  scale_colour_gradient(low = "#deebf7", high = "#08306b", name = "Affective\nScore") +
  facet_wrap(~country) +
  labs(
    title    = "Ha1: Cognitive Beliefs → Purchase Frequency",
    subtitle = "Point shade = Affective Emotion composite; line = OLS fit",
    x = "Cognitive Beliefs (mean score)", y = "Purchase Frequency (1–6)"
  ) +
  theme_study()

# ggsave("Ha1_cognitive_affective_mediation.png", p_ha1,
#        width = 10, height = 5, dpi = 300)




```

## The Theory of the Certification Trust Chain

`Alternative (`H₁`)`: Purchase likelihood is increased by certification familiarity, which then boosts confidence after the purchase.

`Null (``H₀``)`: buy experience does not boost post-buy confidence, and certification familiarity does not raise purchase likelihood.

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-align: "center"


# 7.  Ha2: THEORY OF THE CERTIFICATION TRUST CHAIN
#
#  Path:  Cert Familiarity (C5/C6)
#            → Purchase Frequency (B4)          [a-path]
#            → Post-purchase Confidence (D2)    [b-path]
#  D2 ("I feel confident when I see an organic cert label") is the
#  closest available proxy for post-purchase confidence.


# ── 7.1  Complete-case dataset for Ha2 ────────────────────────────────────────

ha2_data <- df |>
  select(country, cert_eu_n, cert_zw_n, cert_familiar_n,
         cert_imp_n, d2_assurance_emotion_n, purchase_freq_n, f1_intent_n) |>
  drop_na()

cat(sprintf("\nHa2 n = %d\n", nrow(ha2_data)))

# ── 7.2  Baron-Kenny steps ────────────────────────────────────────────────────

# Step 1: total effect — Cert familiarity → Post-purchase confidence (D2)

bk1_ha2 <- lm(d2_assurance_emotion_n ~ cert_eu_n + cert_zw_n, data = ha2_data)
rob_test(bk1_ha2, "Step 1 – Total: Cert familiarity → D2 confidence")

# Step 2: a-path — Cert familiarity → Purchase frequency

bk2_ha2 <- lm(purchase_freq_n ~ cert_eu_n + cert_zw_n, data = ha2_data)
rob_test(bk2_ha2, "Step 2 – a-path: Cert familiarity → Purchase_Freq")

# Step 3: b & c' paths — Cert + Purchase → D2

bk3_ha2 <- lm(d2_assurance_emotion_n ~ cert_eu_n + cert_zw_n + purchase_freq_n,
               data = ha2_data)

rob_test(bk3_ha2, "Step 3 – b & c' paths: Cert + Purchase → D2 confidence")

# ── 7.3  Sobel test ───────────────────────────────────────────────────────────

# Using EU logo as primary treatment (ZW logo held in model as covariate)

sobel_test(bk2_ha2, treat = "cert_eu_n", model_b = bk3_ha2, mediator = "purchase_freq_n")

# ── 7.4  Bootstrapped mediation ───────────────────────────────────────────────

cat("\n--- Ha2: Bootstrapped ACME / ADE / Total (EU logo → Purchase → Confidence) ---\n")

ha2_med <- run_mediation(bk2_ha2, bk3_ha2,
                          treat    = "cert_eu_n",
                          mediator = "purchase_freq_n")

# ── 7.5  SEM ──────────────────────────────────────────────────────────────────

ha2_sem_spec <- '
  purchase_freq_n         ~ a2 * cert_eu_n + a2z * cert_zw_n
  d2_assurance_emotion_n  ~ b2 * purchase_freq_n +
                             c2 * cert_eu_n + c2z * cert_zw_n
  indirect2 := a2 * b2
  total2    := c2 + (a2 * b2)
'

ha2_fit <- sem(ha2_sem_spec, data = ha2_data,
               estimator = "ML", se = "bootstrap", bootstrap = 2000)

sem_fit_summary(ha2_fit, "Ha2 SEM Fit Indices")

cat("\n--- Ha2: Path estimates ---\n")
sem_paths(ha2_fit, c("a2", "b2", "c2", "indirect2", "total2"))

decide_indirect(ha2_fit, "indirect2")

# ── 7.6  Plot ─────────────────────────────────────────────────────────────────

ha2_data |>
  mutate(cert_label = factor(cert_familiar_n,
                             levels = 1:3,
                             labels = c("Unrecognised", "Seen", "Familiar"))) |>

  ggplot(aes(x = cert_label, y = d2_assurance_emotion_n)) +
  geom_boxplot(aes(fill = cert_label), alpha = 0.7, show.legend = FALSE) +
  scale_fill_brewer(palette = "Greens") +
  facet_wrap(~country) +
  labs(
    title    = "Ha2: Certification Familiarity → Post-purchase Confidence (D2)",
    subtitle = "Country-appropriate logo (EU for Poland; ZW for Zimbabwe)",
    x = "Certification Familiarity", y = "D2 Confidence Score (1–5)"
  ) +
  theme_study()

# ggsave("Ha2_cert_trust_chain.png", p_ha2, width = 10, height = 5, dpi = 300)


```

## The Moderating Hypothesis of Price Barriers

`Alternative (H₁)`: The positive correlation between affective sentiments and purchase intentions is weakened by price perception, particularly for consumers with lower incomes.

`Null (H₀)`: Regardless of economic level, the link between affective emotions and purchase intentions is not weakened by price perception.

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-align: "center"

# 8.  Ha3: MODERATING HYPOTHESIS OF PRICE BARRIERS

#  Model:  F1_intent ~ Affective × Price_Perception × Employment_Group
#  Ha3 tests whether high price perception weakens the positive
#  Affective → Intent link, especially among lower-SES consumers.
#  Mean-centred predictors reduce multicollinearity in interaction terms.
# 
ha3_data <- df |> 
  select(country, Affective, Affective_c, price_perception_n, price_c,
         employment_n, f1_intent_n) |>
  drop_na()

cat(sprintf("\nHa3 n = %d\n", nrow(ha3_data)))

# ── 8.2  Base 2-way moderation: Affective × Price → Intent ───────────────────

ha3_base <- lm(f1_intent_n ~ Affective_c * price_c, data = ha3_data)
rob_test(ha3_base, "Ha3 Base: Affective × Price → F1_intent")

cat(sprintf("  R² = %.4f\n", summary(ha3_base)$r.squared))

# ── 8.3  3-way moderation: adds Employment as second moderator ────────────────

ha3_3way <- lm(f1_intent_n ~ Affective_c * price_c * employment_n, data = ha3_data)
rob_test(ha3_3way, "Ha3 3-Way: Affective × Price × Employment → F1_intent")

cat(sprintf("  R² = %.4f  |  ΔR² = %.4f\n",
            summary(ha3_3way)$r.squared,
            summary(ha3_3way)$r.squared - summary(ha3_base)$r.squared))

# ── 8.4  Model comparison via ANOVA ───────────────────────────────────────────

cat("\n--- Ha3: ANOVA — does adding Employment improve fit? ---\n")

print(anova(ha3_base, ha3_3way))

# ── 8.5  Simple slopes by employment tier ─────────────────────────────────────

# Low = 1–2 (part-time / full-time employed at lower scale)

# High = 5–6 (retired / student — proxy for constrained income)

# simple_slopes <- function(emp_vals, label) {
#   m <- lm(f1_intent_n ~ Affective_c * price_c,
#           data = filter(ha3_data, employment_n %in% emp_vals))
#   rob_test(m, paste("Simple slopes —", label))
# }
# 
# simple_slopes(1:2, "Lower employment tier")
# simple_slopes(5:7, "Higher/non-employment tier")

# ── 8.6  Johnson-Neyman interval ─────────────────────────────────────────────

cat("\n--- Ha3: Johnson-Neyman — where is price moderation significant? ---\n")

jn <- interactions::johnson_neyman(
  model = ha3_base, pred = "Affective_c", modx = "price_c", alpha = 0.05
)

print(jn)



# ── 8.8  Plot ─────────────────────────────────────────────────────────────────

# interactions::interact_plot(
#   ha3_3way,
#   pred        = Affective,
#   modx        = price_perception_num,
#   mod2        = employment_n,
#   modx.values = c(-1, 0, 1),
#   x.label     = "Affective Emotions (mean-centred)",
#   y.label     = "Purchase Intention (F1)",
#   legend.main = "Price Barrier\n(±1 SD)",
#   main.title  = "Ha3: Affective × Price × Employment → Purchase Intention") +
#   theme_study()


# ggsave("Ha3_price_moderation.png", p_ha3, width = 11, height = 6, dpi = 300)


```

## **Hypothesis of Country-Specific Demographic Drivers** 

`Alternative (H₁)`: In both nations, socioeconomic position has a favorable impact on cognitive beliefs and purchase frequency, but it does so through distinct mechanisms—directly in Zimbabwe and indirectly through education and urban residency in Poland.

`Null (H₀)`: Neither country's cognitive views nor purchase frequency are significantly impacted by socioeconomic position.

```{r}
#| fig-width: 10
#| fig-height: 8
#| fig-align: "center"

# 9.  Ha4: COUNTRY-SPECIFIC DEMOGRAPHIC DRIVERS
#
#  Hypothesis:
#    Zimbabwe — SES affects Cognitive Beliefs and Purchase Frequency DIRECTLY
#    Poland   — SES operates INDIRECTLY via Education & Urban residency → Cognitive
#
#  SEP composite = mean(household_econ_n, education_n, employment_n)


# ── 9.1  Country subsets ──────────────────────────────────────────────────────

ha4_data <- df |>
  select(country, SEP, household_econ_n, education_n, employment_n,
         Cognitive, purchase_freq_n) |>
   drop_na()


df_pol <- filter(ha4_data, country == "Poland")

df_zim <- filter(ha4_data, country == "Zimbabwe")

cat(sprintf("\nHa4: n(Poland) = %d  |  n(Zimbabwe) = %d\n",
            nrow(df_pol), nrow(df_zim)))

# ── 9.2  OLS by country (baseline robustness check) ──────────────────────────

ols_pol <- lm(purchase_freq_n ~ SEP + education_n + Cognitive, data = df_pol)

ols_zim <- lm(purchase_freq_n ~ SEP + education_n + Cognitive, data = df_zim)

rob_test(ols_pol, 
         "Ha4 OLS — Poland: SEP + Education + Cognitive → Purchase_Freq")

rob_test(ols_zim, 
         "Ha4 OLS — Zimbabwe: SEP + Education + Cognitive → Purchase_Freq")

# ── 9.3  Chow structural-break test ──────────────────────────────────────────

ols_pool <- lm(purchase_freq_n ~ SEP + education_n + Cognitive, data = ha4_data)

rss_pool <- sum(resid(ols_pool)^2)
rss_split <- sum(resid(ols_pol)^2) + sum(resid(ols_zim)^2)
k  <- length(coef(ols_pool))
n_total    <- nrow(ha4_data)
chow_F     <- ((rss_pool - rss_split) / k) / (rss_split / (n_total - 2 * k))
chow_p     <- pf(chow_F, df1 = k, df2 = n_total - 2 * k, lower.tail = FALSE)

cat(sprintf("\n--- Chow Test: F(%d, %d) = %.4f  p = %.6f ---\n",
            k, n_total - 2 * k, chow_F, chow_p))

# ── 9.4  Multi-group SEM (free vs constrained) ────────────────────────────────

# The model encodes both country pathways simultaneously.

# Poland pathway: SEP → Education → Cognitive → Purchase (indirect)

# Zimbabwe pathway: SEP → Cognitive & Purchase directly

ha4_sem_spec <- '
  # Poland mechanism: SEP flows through education to cognitive beliefs
  education_n    ~ m_edu * SEP
  # Cognitive beliefs: driven by SEP and (for Poland) education
  Cognitive      ~ a_sep * SEP + a_edu * education_n
  # Purchase frequency: direct SEP effect + through cognition
  purchase_freq_n ~ b_cog * Cognitive + d_sep * SEP
  # Defined indirect effects
  indirect_sep_cog  := a_sep * b_cog
  indirect_edu_path := m_edu * a_edu * b_cog
  total_indirect    := indirect_sep_cog + indirect_edu_path
  total_sep_effect  := d_sep + total_indirect
'

# Free model: all paths estimated separately per country

ha4_free <- sem(ha4_sem_spec, data = ha4_data, 
                group = "country",
                estimator = "ML", 
                se = "bootstrap", bootstrap = 1000,
                missing = "ml")

# Constrained model: regression paths forced equal across countries (null model)

ha4_const <- sem(ha4_sem_spec, data = ha4_data, group = "country",
                 group.equal = "regressions",
                 estimator = "ML", se = "bootstrap", bootstrap = 1000,
                 missing = "ml")

sem_fit_summary(ha4_free, "Ha4 Free (unconstrained) Model")

cat("\n--- Ha4: Δχ² test — free vs constrained ---\n")

print(lavTestLRT(ha4_free, ha4_const))

# Per-country structural path estimates

pe_free <- parameterEstimates(ha4_free, boot.ci.type = "bca.simple", ci = TRUE) |>
  filter(op %in% c("~", ":=")) |>
  select(lhs, op, rhs, group, est, se, pvalue, ci.lower, ci.upper)

cat("\n--- Ha4: Poland paths (group = 1) ---\n")

print(filter(pe_free, group == 1))

cat("\n--- Ha4: Zimbabwe paths (group = 2) ---\n")

print(filter(pe_free, group == 2))

# ── 9.5  Country-specific mediation confirmatory tests ────────────────────────

# Poland: SEP → Education (mediator) → Cognitive → Purchase

cat("\n--- Ha4 Poland: SES → Education → Cognitive (mediation) ---\n")

pol_m <- lm(education_n ~ SEP, data = df_pol)
pol_y <- lm(Cognitive ~ SEP + education_n, data = df_pol)

ha4_med_pol <- run_mediation(pol_m, pol_y, treat = "SEP", mediator = "education_n",
                              sims = 1000)

# Zimbabwe: SEP → Cognitive (mediator) → Purchase

cat("\n--- Ha4 Zimbabwe: SES → Cognitive → Purchase (mediation) ---\n")

zim_m <- lm(Cognitive  ~ SEP, data = df_zim)
zim_y <- lm(purchase_freq_n ~ SEP + Cognitive, data = df_zim)

ha4_med_zim <- run_mediation(zim_m, zim_y, treat = "SEP", mediator = "Cognitive",
                              sims = 1000)

# ── 9.6  Decision ─────────────────────────────────────────────────────────────

lrt_result <- lavTestLRT(ha4_free, ha4_const)
delta_p    <- lrt_result[["Pr(>Chisq)"]][2]

cat("\n>>> Ha4 Decision:\n")

cat(sprintf("  Chow test: F = %.4f  p = %.4f  →  %s\n",
            chow_F, chow_p,
            if (chow_p < 0.05) "REJECT H\u2080 (structural break confirmed)"
            else "FAIL TO REJECT H\u2080"))


# ── 9.7  Coefficient plot (Poland vs Zimbabwe) ────────────────────────────────

coef_tbl <- bind_rows(
  tidy(ols_pol, conf.int = TRUE) |> mutate(Country = "Poland"),
  tidy(ols_zim, conf.int = TRUE) |> mutate(Country = "Zimbabwe")
) |>
  filter(term != "(Intercept)") |>
  mutate(term = recode(term,
    "SEP"         = "SEP (composite)",
    "education_n" = "Education",
    "Cognitive"   = "Cognitive Beliefs"
  ))

ggplot(coef_tbl,
                aes(x = estimate, y = term, colour = Country,
                    xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_pointrange(position = position_dodge(width = 0.6),
                  linewidth = 0.8, size = 0.9) +
  scale_colour_manual(values = COL_COUNTRY) +
  labs(
    title    = "Ha4: Demographic Drivers of Purchase Frequency by Country",
    subtitle = "OLS coefficients ± 95% CI (HC3 robust)",
    x = "Coefficient estimate", y = NULL
  ) +
  theme_study()

# ggsave("Ha4_country_demographic_drivers.png", p_ha4,
#        width = 9, height = 5, dpi = 300)



```

## Consolidated Models

```{r}
modelsummary(
  list(
    # Ha1
    "Aff ~ Cog"        = bk2,
    "PF ~ Cog + Aff"   = bk3,
    # Ha2
    "PF ~ Cert"        = bk2_ha2,
    "D2 ~ Cert + PF"   = bk3_ha2,
    # Ha3
    "Ha3 Base"         = ha3_base,
    "Ha3 3-Way"        = ha3_3way,
    # Ha4
    "Ha4 Poland"       = ols_pol,
    "Ha4 Zimbabwe"     = ols_zim
  ),
  statistic = "({std.error}) [{conf.low}, {conf.high}]",
  stars     = c("*" = .10, "**" = .05, "***" = .01),
  vcov      = "HC3",
  notes     = "HC3 heteroskedasticity-robust standard errors in parentheses; 95% CIs in brackets.",
  output    = "Hypothesis_Results_Table.docx"
)

```
